---
title: "BadgerNet Analysis"
output: 
  pdf_document:
    toc: true
    number_sections: true
    
knit: (function(inputFile, encoding) {
      out_dir <- "../../../outputs/reports";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

imp_thresh = 3

library(writexl)
library(dplyr)
library(openxlsx)
library(stringr)
library(ggplot2)
library(gtsummary)

source("~/Main work/MiscCode/r-regression-tools/r-regression-tools.R")
setwd("~/Main work/Maternity outcomes/maternity-outcomes/R")

# Load data
badger_load <- load_with_ref(
  "../data/BadgerNet/BadgerNet-processed.parquet",
  seperate_ref_sheet = "../data/BadgerNet/BadgerNet-RefGroups.xlsx",
  ref_sheet = "reference",
  add_colon = FALSE,
  return_ref = TRUE
)

badger <- badger_load[[1]] %>% 
  mutate(
    "Social Services" = SocialServicesInvolvement,
    "LD" = LearningDisabilities,
    "Folic Acid Taken" = FolicAcidTakenDuringPregnancy,
    "Consanguineous Union" = Consanguineous_Relationship,
    "Financial/Housing Issues" = as.factor(case_when(
      HousingIssues =="Yes" ~ "Yes",
      FinancialIssues =="Yes" ~ "Yes",
      TRUE ~ "No",
    )),
    "English Difficulties" = as.factor(case_when(
      DiffUndEnglish == "Yes" ~ "Yes",
      InterpreterRequired == "Yes" ~ "Yes",
      TRUE ~ "NO"
    )),
    "Age Group" = as.factor(case_when(
      Age < 20 ~ "Less than 20",
      Age < 30 ~ "20-29",
      Age < 40 ~ "30-39",
      Age >= 40 ~ "40+",
      TRUE ~ "Unknown"
    ))
    ) 

badger$`Age Group` = factor(badger$`Age Group` , levels=c( "Less than 20", "20-29", "30-39", "40+"))
badger$`Age Group` <- relevel(factor(badger$`Age Group`), ref = "20-29")

# Change reference names for plotting purposes
ref <- badger_load[[2]] 
levels(ref$Variable)[match("SocialServicesInvolvement",levels(ref$Variable))] <- "Social Services"
levels(ref$Variable)[match("LearningDisabilities",levels(ref$Variable))] <- "LD"
levels(ref$Variable)[match("FolicAcidTakenDuringPregnancy",levels(ref$Variable))] <- "Folic Acid Taken"
levels(ref$Variable)[match("Consanguineous_Relationship",levels(ref$Variable))] <- "Consanguineous Union"
levels(ref$Variable)[match("Consanguineous_Relationship",levels(ref$Variable))] <- "Consanguineous Union"

# List all variables of interest
all_variables = c("NumberOfBabies","Age Group",
"AlcoholAbuse","substanceAbuse","Social Services",
"DomesticAbuse","LD","English Difficulties",
"Unsupported","MentalHealth","Homeless","HousingIssues",
"FinancialIssues","sensoryAndPhysicalDis","Citizenship",
"BMI>35","IMD Quintile","Mother Ethnicity",
"SmokingAtDelivery","Gestational_Diabetes",
"Consanguineous Union","Folic Acid Taken", "FGM", 
"Late booking")

# limit reference to variables of interest
ref <- ref %>%
  filter(Variable %in% all_variables)

date1 = min(badger$`MthYr`)
date2 = max(badger$`MthYr`)
time_diff = as.numeric(difftime(max(badger$`MthYr`), min(badger$`MthYr`)))

```

# BadgerNet

The BadgerNet dataset comes from the Birmingham Local Maternity and Neonatal System (LMNS). This dataset contains `r nrow(badger)` unique births from `r date1` to `r date2` with a total timespan of `r time_diff` days. 

## Reference Group

The risk ratios for each of the regressions calculated below are given relative 
to the following reference group. 

```{r comment='', echo=FALSE, results='asis'}
 knitr::kable(ref, caption = "Reference group for logistic regression.")
```

## Variable Correlation Matrix

```{r calc_corr, include=FALSE}

# Convert variables to 1/0 to enable correlation matrix to be calculated
badger2 <- badger %>%
  mutate(
    Unsupported = case_when(
      Unsupported == "Yes" ~ 1,
      TRUE ~ 0
    ),
    `Substance Abuse` = case_when(
      `substanceAbuse` == "Yes" ~ 1,
      TRUE ~ 0
    ),
   `Social Services` = case_when(
     `Social Services` == "Yes" ~ 1,
     TRUE ~ 0
   ),
    `Smoking At Delivery` = case_when(
     `SmokingAtDelivery` == "Yes" ~ 1,
     TRUE ~ 0
     ),
   `Sensory/Physical Disability` = case_when(
     sensoryAndPhysicalDis == "Yes" ~ 1,
     TRUE ~ 0
   ),
   Twins = case_when(
     NumberOfBabies == 2 ~ 1,
     TRUE ~ 0
   ),
    `Mental Health` = case_when(
     `MentalHealth` == "Yes" ~ 1,
     TRUE ~ 0
   ),
   `Learning Disability` = case_when(
     `LD` == "Yes" ~ 1,
     TRUE ~ 0
   ),
   `Late booking` = case_when(
     `Late booking` == "Yes" ~ 1,
     TRUE ~ 0
   ),
   `English Difficulties` = case_when(
     `English Difficulties` == "Yes" ~ 1,
     TRUE ~ 0
   ),
  `Deprived Area` = case_when(
    as.numeric(`IMD Quintile`) %in% c("1", "2") ~ 1,
    TRUE ~ 0),
  `Housing Issues` = case_when(
     `HousingIssues` == "Yes" ~ 1,
     TRUE ~ 0
   ),
  `Gestational Diabetes` = case_when(
     `Gestational_Diabetes` == "Yes" ~ 1,
     TRUE ~ 0
   ),
  `Folic Acid Taken` = case_when(
     `Folic Acid Taken` == "Yes" ~ 1,
     TRUE ~ 0
   ),
  `Financial Issues` = case_when(
     `FinancialIssues` == "Yes" ~ 1,
     TRUE ~ 0
   ),
   FGM = case_when(
     `FGM` != "Yes" ~ 1,
     TRUE ~ 0),
   `Global Majority Ethnicity` = case_when(
     `Ethnicity Group` != "White" ~ 1,
     TRUE ~ 0
   ),
  `Domestic Abuse` = case_when(
     `DomesticAbuse` == "Yes" ~ 1,
     TRUE ~ 0
   ),
  
  `Consanguineous Union` = case_when(
     `Consanguineous Union` == "Yes" ~ 1,
     TRUE ~ 0
   ),
  
  Citizenship = case_when(
     `Citizenship` == "Yes" ~ 1,
     TRUE ~ 0
   ),
  
  `BMI>35` = case_when(
     `BMI>35` == "Yes" ~ 1,
     TRUE ~ 0
     ),

   `Alcohol Abuse` = case_when(
     `AlcoholAbuse` == "Yes" ~ 1,
     TRUE ~ 0
   ),

  `Age < 20` = case_when(
     Agelessthan20 == "Yes" ~ 1,
     TRUE ~ 0
   )
   ) %>%
  select(c(Unsupported, `Substance Abuse`, `Social Services`, `Smoking At Delivery`, `Sensory/Physical Disability`, Twins, `Mental Health`, `Learning Disability`, `Late booking`, `Deprived Area`, `Housing Issues`, `Gestational Diabetes`, `Folic Acid Taken`, `Financial Issues`, FGM, `Global Majority Ethnicity`, `Domestic Abuse`, `English Difficulties`, `Consanguineous Union`, Citizenship, `BMI>35`, `Alcohol Abuse`, `Age < 20`
           )) %>%
  tidyr::drop_na()
```

The following is a correlation matrix for our primary variables.

```{r corr_mat}
pdf(file = "../outputs/figures/correlation_2.pdf")
par(family="serif")
corrplot::corrplot(cor(badger2),
         method="square", 
         diag = FALSE, 
         addgrid.col = TRUE,
         type = 'lower',
         tl.col="black")
dev.off()
```

```{r Fix Variable Names}
badger <- badger %>%
  mutate(
    Twins = case_when(
      NumberOfBabies == 2 ~ "Yes",
      TRUE ~ "No"
      ),
    `Domestic Abuse` = DomesticAbuse,
    `Mental Health Issue(s)` = MentalHealth,
    `Financial Issue(s)` = FinancialIssues,
    `Housing Issue(s)` = HousingIssues,
    `Sensory/Physical Disability` = sensoryAndPhysicalDis,
    `Alcohol Abuse` = AlcoholAbuse,
    `Substance Abuse` = substanceAbuse,
    `Smoking at Delivery` = SmokingAtDelivery,
    `Gestational Diabetes` = Gestational_Diabetes
    )

badger$Twins <- relevel(factor(badger$Twins), ref = "No")

all_variables = c("Twins","Age Group",
"Alcohol Abuse","Substance Abuse","Social Services",
"Domestic Abuse","LD","English Difficulties",
"Unsupported","Mental Health Issue(s)","Homeless","Housing Issue(s)",
"Financial Issue(s)","Sensory/Physical Disability","Citizenship",
"BMI>35","IMD Quintile","Mother Ethnicity",
"Smoking at Delivery","Gestational Diabetes",
"Consanguineous Union","Folic Acid Taken", "FGM", 
"Late booking")

```


# Basic regression

First we perform a basic regression with all variables of interest.

## Low-birth weight
```{r basic_reg_lbw, echo=FALSE}
# build formula
lbw_formula <- write_formula("`LowBirthWeight`", all_variables, quote_wrap = TRUE)
print(lbw_formula)

# run logistic regression
model_lbw <- glm(lbw_formula, family = binomial, data=badger)

# find most important variables
varImp_lbw <- var_imp(model_lbw)
head(varImp_lbw, 10)

```


```{r lbw_vif, echo=FALSE}
#create vector of VIF values
vif_values <- car::vif(model_lbw)

lbw_vif <- data.frame(
  Variable = row.names(vif_values),
  VIF = as.numeric(vif_values[,1]),
  Outcome = "LBW"
  )

```


## Premature birth
```{r basic_reg_prem, echo=FALSE}
# build formula
prem_formula <- write_formula("`Premature`", all_variables, quote_wrap = TRUE)
print(prem_formula)

# run logistic regression
model_prem <- glm(prem_formula, family = binomial, data=badger)

# find most important variables
varImp_prem <- var_imp(model_prem)
head(varImp_prem, 10)
```

```{r prem_vif, echo=FALSE}
#create vector of VIF values
vif_values <- car::vif(model_prem)

prem_vif <- data.frame(
  Variable = row.names(vif_values),
  VIF = as.numeric(vif_values[,1]),
  Outcome = "Premature"
  )

```

## Stillbirth

```{r basic_reg_stillbirth, echo=FALSE}
# build formula
sb_formula <- write_formula("`StillBirth`", all_variables, quote_wrap = TRUE)
print(sb_formula)

# run logistic regression
model_sb <- glm(sb_formula, family = binomial, data=badger)

# find most important variables
varImp_sb <- var_imp(model_sb)
head(varImp_sb, 10)
```

```{r sb_vif, echo=FALSE}
#create vector of VIF values
vif_values <- car::vif(model_sb)

sb_vif <- data.frame(
  Variable = row.names(vif_values),
  VIF = as.numeric(vif_values[,1]),
  Outcome = "Stillbirth"
  )

```

## Neonatal death

```{r basic_reg_NeonatalDeath, echo=FALSE}
# build formula
nd_formula <- write_formula("`NeonatalDeath`", all_variables, quote_wrap = TRUE)
print(nd_formula)

# run logistic regression
model_nd <- glm(nd_formula, family = binomial, data=badger)

# find most important variables
varImp_nd <- var_imp(model_nd)
head(varImp_nd, 10)
```

```{r nd_vif, echo=FALSE}
#create vector of VIF values
vif_values <- car::vif(model_nd)

nd_vif <- data.frame(
  Variable = row.names(vif_values),
  VIF = as.numeric(vif_values[,1]),
  Outcome = "NeonatalDeath"
  )

```

## VIF

```{r vif, echo=FALSE}

library(elementalist) # 

#combine VIF values

all_vif <- rbind(lbw_vif, prem_vif, sb_vif, nd_vif)

# Colour blind friendly pallet
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

all_vif$Variable <- gsub("`", "",all_vif$Variable)


all_vif$Variable <- stringr::str_wrap(all_vif$Variable, width = 13)

vif_plot <- ggplot(all_vif, aes(y = VIF, x = Variable, fill = Outcome)) +
  geom_bar(stat="identity", position = "dodge") +
  scale_fill_manual(values=cbPalette) +
  #geom_hline(yintercept = 5, linetype='dashed') +
  theme_bw() +
  theme(
    legend.position.inside = c(0.9, 0.7),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    text = element_text(family="serif", size = 10, colour = "black"),
    legend.text = element_text(size=7),
    legend.key = element_rect(fill = "transparent"),
    legend.background = element_rect(fill='transparent'), 
    legend.box.background = element_rect_round(radius = unit(0.1, "snpc"),fill='white',colour = 'gray'),
    legend.key.height= unit(5, 'mm'),
    legend.key.width= unit(5, 'mm')
  ) +
  labs(fill = NULL) +
  ylim(0, 4) +
  xlab("") +
  ylab("Variance Inflation Factor") 

vif_plot

ggsave("../outputs/figures/vif.pdf", vif_plot,
       height = 3.5, width = 8)

```

## Basic regression final results
```{r basic_reg_vis, echo=FALSE, warning=FALSE}
# merge tables
tbl_output_intermediate <-
  tbl_merge(
    tbls = list(
      tbl_regression(model_lbw, exponentiate = TRUE),
      tbl_regression(model_prem, exponentiate = TRUE)
      ),
    tab_spanner = c("**Low birth weight**", "**Premature Birth**")
  ) %>%
  as_gt()

tbl_output_intermediate %>%
  gt::gtsave(filename = "../../../outputs/BN_int_reg_table.tex")

# Print out table
tbl_output_intermediate

# merge tables
tbl_output_final <-
  tbl_merge(
    tbls = list(
      tbl_regression(model_sb, exponentiate = TRUE),
      tbl_regression(model_nd, exponentiate = TRUE)
      ),
    tab_spanner = c("**Stillbirth**", "Neonatal Death")
  ) %>%
  as_gt()

tbl_output_final %>%
  gt::gtsave(filename = "../../../outputs/BN_reg_table.tex")

# Print out table
tbl_output_final
```

We see that the result is not statistically significant for the following variables: AgeLessThan20, AlcoholAbuse, LD, DiffUndEnglish, Unsupported, Homeless, Citizenship, InterpreterRequired, FGM. Therefore, we remove these variables from our model.

```{r add_marginal, include=FALSE}

all_important <- c(
  "Mother Ethnicity", "IMD Quintile", "Financial/Housing Issues",
  "Substance Abuse","Social Services", "Age Group",
  "Twins", "Mental Health Issue(s)","Sensory/Physical Disability",
  "BMI>35", "Gestational Diabetes", "Smoking at Delivery",
  "Folic Acid Taken", "Late booking", "> 4 missed apts",
  "Consanguineous Union")

formula_plus <- write_formula("`Output:`", all_important, quote_wrap = TRUE)


less_variables <- c("Ethnicity Group","IMD Quintile", "Financial/Housing Issues",
                    "Age Group",   "Twins", "Mental Health Issue(s)", 
                  "Smoking at Delivery","Folic Acid Taken", "Late booking"
           )

formula_minus <- write_formula("`Output:`", less_variables, quote_wrap = TRUE)
```

Our maximised formula now becomes:

`r formula_plus`

While our reduced formula is instead:

`r formula_minus`

# Population description

First, lets calculate the proportions of each outcome for each variable

```{r get_ref_summary_int, results='asis'}
table1 <- badger %>%
  mutate(Outcome = case_when(
    Premature ~ "Premature",
    LowBirthWeight ~ "LBW",
    TRUE ~ "Normal"
  ),
  Outcome = factor(Outcome, levels = c("Normal", "LBW", "Premature"))
  ) %>%
  tbl_summary(
    include = all_important,
    by = "Outcome", # split table by group
  ) %>%
  #add_overall() %>% 
  #add_ci() %>%
  modify_header(label = "**Variable**") %>% # update the column header
  bold_labels() 

table1

table1 %>%
   as_gt() %>%
   gt::gtsave(filename = "../outputs/BN_int_reg_table.tex")
```

```{r get_ref_summary_final, results='asis'}
table2 <- badger %>%
  mutate(Outcome = case_when(
    StillBirth == 1 ~ "Stillbirth",
    NeonatalDeath == 1 ~ "Neonatal Death",
    TRUE ~ "Normal"
  ),
  Outcome = factor(Outcome, levels = c("Normal", "Stillbirth", "Neonatal Death"))
  ) %>%
  tbl_summary(
    include = less_variables,
    by = "Outcome", # split table by group
  ) %>%
  modify_header(label = "**Variable**") %>% # update the column header
  bold_labels() 

table2

table2 %>%
   as_gt() %>%
   gt::gtsave(filename = "../outputs/BN_final_reg_table.tex")
```

# Final Regression

## Rerunning with fewer variables

```{r final_reg, include = FALSE}

# Maximised formula 
lbw_ff <- write_formula("LowBirthWeight", all_important, quote_wrap = TRUE)
model_lbw_final <- glm(lbw_ff, family = binomial, data=badger)

prem_ff <- write_formula("Premature", all_important, quote_wrap = TRUE)
model_prem_final <- glm(prem_ff, family = binomial, data=badger)

# Reduced formula
nd_ff <- write_formula("NeonatalDeath", less_variables, quote_wrap = TRUE)
model_nd_final <- glm(nd_ff, family = binomial, data=badger)

sb_ff <- write_formula("StillBirth", less_variables, quote_wrap = TRUE)
model_sb_final <- glm(sb_ff, family = binomial, data=badger)

```

## Display results

```{r merge_tables, echo=FALSE, warning=FALSE}
set.seed(1)

lbw_exp = tbl_regression(model_lbw_final, exponentiate = TRUE) 
prem_exp = tbl_regression(model_prem_final, exponentiate = TRUE)
sb_exp = tbl_regression(model_sb_final, exponentiate = TRUE)
nd_exp = tbl_regression(model_nd_final, exponentiate = TRUE)
```

```{r Final Results save, include=FALSE}

lbw_exp = lbw_exp %>% 
  as_tibble() 
  
prem_exp = prem_exp %>% 
  as_tibble() 

sb_exp = sb_exp %>% 
  as_tibble() 

nd_exp = nd_exp %>%
  as_tibble()

dataset_names <- list('LBW' = lbw_exp, 
                      'Premature' = prem_exp,
                      'StillBirth' = sb_exp,
                      'NeonatalDeath' = nd_exp)

openxlsx::write.xlsx(dataset_names, 
                     file = '../data/BadgerNet/BadgerNet-reg-results.xlsx') 
```


```{r forest_plot_function}
library(forestplot)

fplot <- function(model_tib, 
                  title = "",
                  part = 1) {

  
  # model_tib = tibble of the regression table
  model_tib["**95% CI** 2"] <- model_tib["**95% CI**"]
  model_tib <- tidyr::separate(model_tib, col = `**95% CI** 2`, into = c("lowerCI", "upperCI"), sep = ",")
  model_tib$`OR [95% CI]` = paste(
      model_tib$`**OR**`, 
      " [", model_tib$lowerCI, 
      "; ", model_tib$upperCI, "]",
      sep = ""
      )
  
  model_tib <- model_tib %>%
  mutate(
    `Odds Label`= case_when(
      is.na(`**OR**`) & 
        `**Characteristic**` %in% colnames(badger) ~ NA,
      is.na(`**OR**`) ~ "Ref",
      TRUE ~ as.character(`**OR**`)
    ),
    `Odds Ratio`= case_when(
      is.na(`**OR**`) & 
        `**Characteristic**` %in% colnames(badger) ~ NA,
      is.na(`**OR**`) ~ 1,
      as.numeric(`**OR**`) > 10 ~ NA,
      TRUE ~ as.numeric(`**OR**`)
    ),
    
    `lowerCI`= case_when(
      is.na(`lowerCI`) & 
        `**Characteristic**` %in% colnames(badger) ~ NA,
      is.na(`lowerCI`) ~ 1,
      TRUE ~ as.numeric(`lowerCI`)
    ),
    `upperCI`= case_when(
      is.na(`upperCI`) & 
        `**Characteristic**` %in% colnames(badger) ~ NA,
      is.na(`upperCI`) ~ 1,
      TRUE ~ as.numeric(`upperCI`)
    ),
    `OR [95% CI]` = case_when(
      is.na(`**OR**`) & 
        `**Characteristic**` %in% colnames(badger) ~ NA,
      is.na(`**OR**`) ~ "Reference",
      TRUE ~ `OR [95% CI]`
    )
  ) 
  

  if (part == 1) {
    # remove characteristic column

    model_tib_reduced <- model_tib %>% 
      select(c("**Characteristic**", "OR [95% CI]", "**p-value**"))
  } else {
    # Do nothing
    model_tib_reduced <- model_tib %>% 
      select(c("OR [95% CI]", "**p-value**"))
  } 
  
  # Prepare forest plot
  new_fp <- model_tib_reduced |> 
    forestplot(
    clip = c(-1, 6),
    graph.pos= 3 - part,
    mean= model_tib["Odds Ratio"], 
    zero=1,
    title=title,
    lower = model_tib["lowerCI"],
    upper = model_tib["upperCI"],
    boxsize=0.2,
    lwd.ci = 1.7,
    line.margin = unit(1, "mm"),
    lineheight = unit(6, "mm"),
    colgap = unit(5, "mm"),
    fn.ci_norm="fpDrawNormalCI",
    graphwidth=unit(0.2, "npc"),
    xticks = c(-1, 0,1, 2,3, 4,5, 6),
    is.summary = model_tib$`**Characteristic**` %in% colnames(badger),
    col=fpColors(box="#1f77b4", lines="#73B7E5", zero = "#D0D0D0"),
    txt_gp=fpTxtGp(label=gpar(cex=0.9),
                ticks=gpar(cex=0.8),
                xlab = gpar(cex=0.9),
                title=gpar(cex=1)),
    ) 
  if (part == 1) {

    new_fp <- new_fp |>
    fp_add_header(
      `**Characteristic**` = c(""),
      `OR [95% CI]` = c("OR [95% CI]"),
      `**p-value**` = c("p-value")
      ) |>
    fp_set_style(
      txt_gp = fpTxtGp(label = gpar(fontfamily = "serif")),
      )
  } else if (part == 2) {
    
    new_fp <- new_fp |>
    fp_add_header(
      `OR [95% CI]` = c("OR [95% CI]"),
      `**p-value**` = c("p-value")
      ) |>
    fp_set_style(
      txt_gp = fpTxtGp(label = gpar(fontfamily = "serif")),
      )
  }
  if (part == 1) {
      new_fp <- new_fp |>
    fp_set_style(align = "lccc")
  } else{
      new_fp <- new_fp |>
    fp_set_style(align = "ccc")
  }


  return(new_fp)
}

```



```{r combine_forest_large}

lbw_fp <- fplot(lbw_exp, title = "Low Birth Weight", part = 1)
prem_fp <- fplot(prem_exp, title = "Premature Birth", part = 2)

p1 <- ggplotify::grid2grob(print(lbw_fp)) 
p2 <- ggplotify::grid2grob(print(prem_fp))
both_big <- cowplot::plot_grid(
  p1,
  NA,
  p2,
  ncol = 3,
  rel_widths = c(1, -0.3, 1)
)
both_big

ggplot2::ggsave(both_big, 
                filename = "../outputs/figures/regression/reg_lbw_prem.pdf", 
                width = 12,
                height = 14)
```

```{r combine_forest_small}

sb_fp <- fplot(sb_exp, title = "Stillbirth", part = 1)
nd_fp <- fplot(nd_exp, title = "Neonatal Death", part = 2)

p1 <- ggplotify::grid2grob(print(sb_fp))
p2 <- ggplotify::grid2grob(print(nd_fp))

both_smalls <- cowplot::plot_grid(
  p1,
  NULL,
  p2,
  ncol = 3,
  rel_widths = c(1, -0.3, 1)
)
both_smalls

ggplot2::ggsave(both_smalls, 
                filename = "../outputs/figures/regression/reg_sb_nd.pdf", 
                width = 12,
                height = 7.8)
```